---
title: "EDA_Practice"
format: html
editor: visual
toc: true
---

```{r}
library(tidyverse)
```

##Task 1: Read in Data and Modify

###Part 1: Read in Data
```{r}
#Provided code from documentation
d1=read.table("student-mat.csv",sep=";",header=TRUE)
d2=read.table("student-por.csv",sep=";",header=TRUE)

d3=merge(d1,d2,by=c("school","sex","age","address","famsize","Pstatus","Medu","Fedu","Mjob","Fjob","reason","nursery","internet"))
print(nrow(d3)) # 382 students
```

###Part 2: Inspecting the issue when using inner join with the selected variables
```{r}
d1 <- read_delim("student-mat.csv", delim = ";", show_col_types = FALSE)
d2 <- read_delim("student-por.csv", delim = ";", show_col_types = FALSE)

d3 <- inner_join(d1, d2, by = c("school","sex","age","address","famsize","Pstatus","Medu","Fedu","Mjob","Fjob","reason","nursery","internet"))
```
*The issue is detected on Row 79. It duplicated columns (G1, G2, G3, etc...) and add a suffix with what data set it came from (x or y). This is happening because it I believe some of the students have overlapping key variables so it a many-to-many join and not a one to one join.*

###Part 3: Use an inner_join() on all variables other than G1, G2, G3, paid, and absences. Use his form of the combined data in further exercises.
```{r}
d1 <- read_delim("student-mat.csv", delim = ";", show_col_types = FALSE)
d2 <- read_delim("student-por.csv", delim = ";", show_col_types = FALSE)

d3 <- inner_join(d1, d2, by = c("school","sex","age","address","famsize","Pstatus","Medu","Fedu","Mjob","Fjob", "reason", "guardian", "traveltime", "studytime", "failures", "schoolsup", "famsup", "activities", "nursery", "higher", "internet", "romantic", "famrel", "freetime", "goout", "Dalc", "Walc", "health"))
```

###Part 4: Next, for the math data, Portuguese, and combined data, choose four categorical variables you are interested in and convert those into factor variables in each tibble (use the same four variables in each). Use the mutate() function to accomplish this.
```{r}
d1 <- d1 |> 
  mutate(Pstatus = as.factor(Pstatus),
         famsize = as.factor(famsize),
         Mjob = as.factor(Mjob),
         Fjob = as.factor(Fjob))

d2 <- d2 |> 
  mutate(Pstatus = as.factor(Pstatus),
         famsize = as.factor(famsize),
         Mjob = as.factor(Mjob),
         Fjob = as.factor(Fjob))

d3 <- d3 |> 
  mutate(Pstatus = as.factor(Pstatus),
         famsize = as.factor(famsize),
         Mjob = as.factor(Mjob),
         Fjob = as.factor(Fjob))

```


##Summarize the Data (Explore Data Set, Categorical Variable Summaries, Numerical Variable Summaries)


**Checking the data set for any missing values**
```{r}
#Looking at the combined data set
str(d3)

#Documenting missing values
summary(d3)
```
*There are no missing values*


###Categorical Variable Summaries


####Creating a one-way, two-way, and three-way contingency table for the factor variables created above\*\*
```{r}
#One-Way contingency table
one_way <- table(d3$famsize)
one_way
```
*There are 230 observations with a family size of more than 3 people and 90 observations with with a family size of 3 or less people*

```{r}
#Two-Way contingency table
two_way <- table(d3$famsize, d3$Mjob)
two_way
```
*Of the families with three or more members, 35 of the male parents work as teachers*

```{r}
three_way <- table(d3$famsize, d3$Mjob, d3$Fjob)
three_way
```
*There are 3 students who have a family size larger than three and who's parents both work at home.*

####Creating a conditional two-way contingency table (2 separate methods)
```{r}
#Filtering data and then creating the table
d3_filt <- d3 |>
  filter(Mjob == "teacher")

method_a <- table(d3_filt$famsize, d3_filt$Fjob)
method_a

#Creating a three-way contingency table and then slicing it
d3_slice <- table(d3$famsize, d3$Fjob, d3$Mjob)

method_b <- d3_slice[ , , "teacher"]
method_b

```
*There are 2 students that a family with more than 3 members in it and have a male parent who works as a teacher as well as female parent who works have home.*

####Creating a two-way contingency table using group_by and summarize. Then using pivot_wider() to make the result look more like the output above\*\*
```{r}
method_c <- d3 |>
  group_by(famsize, Fjob) |>
  summarise(count = n()) |>
  pivot_wider (
    names_from = Fjob,
    values_from = count)

method_c
```
*There are 4 students who have a family size less than or equal to three who female parent works in the health industry.*

####Creating a stacked bar graph using using family size and the female parents job
```{r}
bar_plot <- ggplot(data = d3, aes(x = famsize, fill = Fjob))
bar_plot + geom_bar() +
  labs(title = "Mom's Occupation by Family Size of Students",
       x = "Family Size",
       y= "Number of Students")
```
*The stacked bar plot is a visual representation of the distribution of student's mother's occupation by their family size (either greater than 3 members or less than or equal to 3 members).*


###Numeric Variable Summaries


####Find measures of center and spread for three of these variables (including at least one G3 variable)
```{r}

```

